---
title: "NOAAPackage"
author: PG
date: "2020-04-11"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NOAAPackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The NOAA package contains data on destructive earthquakes from 2150 B.C. to the present provided by the National Centers For Environmental Information on [https://www.ngdc.noaa.gov](https://www.ngdc.noaa.gov).

The package contains a function to read the data and another one to clean to clean it. Moreover, two geoms are available to visualize and label earthquakes on a time line, respectively, as well as two functions to display and label eathquakes on a map, respectively.

First, let's load the necessary packages:

```{r, message=FALSE, warning=FALSE}
library(NOAAPackage)
library(dplyr)
library(stringr)
library(ggplot2)
library(lubridate)
```

Data can be read into R using the function *noaa_read()*:

```{r}
noaa_data <- noaa_read(system.file("extdata",
                                   "NOAA_Significant_Earthquake_Database.txt",
                                   package = "NOAAPackage"))
```

The data set needs to be cleaned using the function *eq_clean_data()* so that the data can later be used with the provided geoms and functions for visualizations and map plots.

```{r, message=FALSE, warning=FALSE}
noaa_data_cleaned <- noaa_data %>% 
  eq_clean_data()
```

Let's have a look at the geom called *geom_timeline()* and at the plot it makes:

```{r, fig.height=4, fig.width=7}
usa_data <- noaa_data_cleaned %>%
  filter(str_trim(COUNTRY) %in% c("USA") & DATE %in% c(ymd("2000-01-01"):ymd("2017-01-01")))

ggplot(data = usa_data) +
    geom_timeline(aes(x = DATE, color = DEATHS, size=EQ_PRIMARY), alpha=0.6)
```

If you want to add lables to the n_max largest earthquakes, you can add another geom called *geom_timeline_label()* after *geom_timeline()*:

```{r, fig.height=6, fig.width=7}
ggplot(data = usa_data) +
    geom_timeline(aes(x = DATE, color = DEATHS, size=EQ_PRIMARY), alpha=0.6) + 
    geom_timeline_label(aes(label = LOCATION_NAME, x = DATE, size=EQ_PRIMARY, n_max = 4)) +
    guides(size = guide_legend(title = "Richter scale value")) +
    scale_colour_continuous(name = "# DEATHS") +
    theme_classic() +
    theme(legend.position="bottom")
```

Earthquakes can also be shown on a map using the NOAA package functions for data cleaning, i.e. *eq_clean_data()* as well as the function to display the earthquake locations on a map, i.e. *eq_map()*. If you click on one of the locations, the date of the earthquake pops up.

```{r, fig.height=6, fig.width=7}
noaa_data %>%
    eq_clean_data() %>%
    dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
    eq_map(annot_col = "DATE")
```

To receive more information about the event such as location, magnitude and total deaths - in case the information is available - you can use the function *eq_create_label()* as shown in the following example:

```{r, fig.height=6, fig.width=7}
noaa_data %>%
    eq_clean_data() %>% 
    dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
    dplyr::mutate(popup_text = eq_create_label(.)) %>% 
    eq_map(annot_col = "popup_text")

```

